Open Linked Social Data for Scientific Benchmarking

### 2016-07-01 paper.tex
advances in written parts

\usepackage{hyperref}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{float}
\usepackage{amssymb}
\usepackage{cleveref}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
%
\pgfplotsset{compat=1.5}
\pgfplotsset
{
	width=0.5\textwidth,
	x tick label style={/pgf/number format/1000 sep=},
  enlarge x limits = 0.0,
  ymajorgrids=true,
	major tick style={draw=none},
  ymin = 0.0,
	every axis/.append style={
		every x tick label/.append style={font=\tiny},
    every y tick label/.append style={font=\tiny},
    every axis label/.append style={font=\small},
    height=37mm,
    width=37mm,
    title style={at={(0.5,0.90)}, font=\normalfont},
    xticklabel style={yshift=4pt}
	}
}

@@

\section{Introduction}
The enormity of the digital data propels a rapid development of analysis methods
from different perspectives.

\subsection{Benchmarking in the analysis of complexity}
%
Karate club, whatelse?

\section{Materials: data from diverse provenance}\label{materials}

\section{Methods}

\subsection{Linked open data}
\subsection{RDF}
\subsection{Data-driven ontology synthesis}

>>>>

\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{float}
\usepackage{amssymb}
\usepackage{cleveref}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
%
\pgfplotsset{compat=1.5}
\pgfplotsset
{
	width=0.5\textwidth,
	x tick label style={/pgf/number format/1000 sep=},
  enlarge x limits = 0.0,
  ymajorgrids=true,
	major tick style={draw=none},
  ymin = 0.0,
	every axis/.append style={
		every x tick label/.append style={font=\tiny},
    every y tick label/.append style={font=\tiny},
    every axis label/.append style={font=\small},
    height=37mm,
    width=37mm,
    title style={at={(0.5,0.90)}, font=\normalfont},
    xticklabel style={yshift=4pt}
	}
}

@@

\section{Introduction}
In recent years, the web of linked data~\cite{lee1} has attracted wide attention in
both research and application realms.
However, there is is a lack of datasets for research benchmarking,
specially in the complex networks field, yielding diverse results from 
poorly related data.

The enormity of the digital data propels a rapid development of analysis methods
from different perspectives.
The used datasets differ within the scope of each research with
scarce and historical exception such as the
karate club dataset~\cite{newmanBook}.

This work presents a linked data dataset with data from diverse
provenance, including Facebook, Twitter, IRC, Email and detached
instances.
Such data is proposed as a common repertoire for scientific
reseach involving networks and textual content.

% \subsection{Benchmarking in the analysis of complexity}

\section{Materials}\label{materials}
\subsection{Facebook data}
\subsection{Twitter data}
\subsection{IRC data}
\subsection{Email data}
\subsection{ParticipaBR data}
\subsection{AA data}
\subsection{Cidade Democr\'atica data}

\section{Methods}
\subsection{Linked open data}
Linked data refers to data published in the web in such a way that it is
machine readable and conforms to a set of best practices.
The web of data is constructed with documents on the web 
such as the web of hypertext.
In practice, the idea of linked data can me summarized
by 1) the use of RDF to publish data on the web and 2) the use of RDF
links to interlink data from different sources.
The web is expected to be interconnected and to grow by the systematic application of four
steps~\cite{lee1}:
\begin{itemize}
    \item Use URIs to identify things~\cite{uri}.
    \item Use HTTP URIs.
    \item Provide useful information when an URI is accessed via HTTP.
    \item Provide other URIs in the description of resources so human
        and machine agents can perform discovery.
\end{itemize}

The Linked Open Data~\cite{lod} builds an ever growing cloud of data,
the global data space, which is usually
conceived as centered around the DBPedia, a linked data representation
of data from Wikipedia~\cite{dbpedia0,dbpedia}.

\subsection{RDF}
The Resource Description Framework (RDF), a W3C
recommendation, is a model for data
interchange.
It is based on the idea of making statements about resources in the form
of triples, i.e. expressions in the form ``subject - predicate -
object''.
RDF can be serialized in several file formats, including RDF/XML,
Turtle and Manchester all which, in essence, represent a labeled and
directed multi-graph.
RDF may be stored in a type of database called a triplestore~\cite{rdf}.

As an example of an RDF statement, the following triple in the Turtle
format asserts that ``the paper has color white'':\\
\texttt{http://example.org/paper http://example.org/hasColor\\
http://example.org/White .}

\subsection{Data-driven ontology synthesis}
OWL Ontologies are critical tools to describe taxonomies and the
structure of knowledge.
Most ontologies are created by domain experts even though the data they
arrange is often given by a software system.

We developed an ontology synthesis method that probes the ontological structure in data with
SPARQL queries and post-processing which can be divided in the following steps:
\begin{enumerate}[leftmargin=0cm]
    \item Obtaining all distinct classes with the query:\\
        \texttt{SELECT DISTINCT ?class WHERE \{ ?s a ?class \}}
    \item Obtaining all distinct properties with the query:\\
        \texttt{SELECT DISTINCT ?p WHERE \{ ?s ?p ?o \}}
    \item For each class, get distinct subject classes and predicates where the
        the object is an instance of the class:\\
        \texttt{SELECT DISTINCT ?p ?cs WHERE \{ ?i a <class\_uri> . ?s ?p ?i . ?s
        a ?cs . \}}
    \item For each class, get distinct predicates and object classes or
        datatypes where the subject is an instance of such class:\\
        \texttt{SELECT DISTINCT ?p ?co (datatype(?o) as ?do) WHERE \{ ?i
                a <class\_uri> . ?i ?p ?o . OPTIONAL \{ ?o a ?co . \} \}}
    \item For each property, check if it is functional, i.e. if it
        occurs only once with each subject:\\
        \texttt{SELECT DISTINCT (COUNT(?o) as ?co) WHERE \{ ?s
            <property\_uri> ?o \} GROUP BY ?s}
    \item For each property, find the incident range and domain with the
        queries:\\
        \texttt{SELECT DISTINCT ?co (datatype(?o) as ?do) WHERE \{ ?s
                <property\_uri> ?o . OPTIONAL \{ ?o a ?co . \} \}}
        \texttt{SELECT DISTINCT ?cs WHERE \{ ?s <property\_uri> ?o . ?s a ?cs . \}}
    \item For each instance of each class, get all distinct predicates.
        For each predicate, check if all instances of the class
        hold such relationship (existential restriction):\\
        \texttt{SELECT DISTINCT ?p WHERE \{ ?s a <class\_uri>. ?s ?p ?o
        . \}}\\
        \texttt{SELECT DISTINCT ?s WHERE \{ ?s a <class\_uri> \}}\\
        \texttt{SELECT DISTINCT ?s ?co  (datatype(?o) as ?do) WHERE \{?s
                a <class\_uri>. ?s <property\_uri> ?o . OPTIONAL \{?o a ?co . \}\}}
    \item and if all instances that hold such relationship are instances of the class
        (universal restriction):\\
        \texttt{SELECT DISTINCT ?s WHERE \{ ?s <property\_uri> ?o . \}}
    \item Draw each class, each property and the overall figure.
    \item Make \texttt{rdfs:subClassOf} and \texttt{rdfs:subPropertyOf}
        statements to better organize knowledge and link to third party
        ontologies and data.
    \item
\end{enumerate}
